---
title: "Metabolomics"
author: "wendyxwang"
date: "Sunday, May 24, 2015"
output:
  ioslides_presentation:
    fig_caption: yes
    fig_retina: 1
    keep_md: yes
    smaller: yes
  beamer_presentation: default
---

```{r,cache=FALSE}
# Set some global knitr options
library("knitr")
opts_chunk$set(tidy=TRUE, size='tiny',tidy.opts=list(blank=FALSE, width.cutoff=80), cache=TRUE, messages=FALSE)
```

## import all the data
```{r}
setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")

#read in subject demographics
library(xlsx)
demographics = read.xlsx("Subjects.xlsx",1)
colnames(demographics) = c("ID","baseline","order","age","sex","weight","height","bmi","bodyfat","notes")
summary(demographics)

#read in sample position and glucose
sample = read.csv("Glance_global_key_samples.csv", header=T, sep=",")
colnames(sample)
sample = sample[,-4]

#read in negative results
neg = read.csv("Neg.csv", header=T, sep=",", na.strings=".")
dim(neg)
sum(is.na(neg))
#impute all NA's with 1
neg[is.na(neg)] = 1
sum(neg == 1)

#read in positive results
pos = read.csv("Pos.csv", header=T, sep=",", na.strings=".")
dim(pos)
sum(is.na(pos))
pos[is.na(pos)] = 1
sum(pos == 1)

#Save RData
save.image(file="Glance.RData")
```


##Combine the dataset
```{r}
colnames(sample)
sample = sample[,1:5]
sample.QC = sample[!is.na(sample[,1]),]
dim(sample.QC)
neg$ID = sample.QC$id
pos$ID = sample.QC$id

sample.QC = sample.QC[which(substr(sample.QC$id,1,2)!="QC"),]
neg.QC = neg[which(substr(sample.QC$id,1,2)!="QC"),]
pos.QC = pos[which(substr(sample.QC$id,1,2)!="QC"),]

#exclude ill person
sample.QC = sample.QC[which(sample.QC$id!="80013"),]
neg.QC = neg.QC[which(neg.QC$ID!="80013"),]
pos.QC = pos.QC[which(pos.QC$ID!="80013"),]

#exclude drop-out
sample.QC = sample.QC[which(sample.QC$id!="80025"),]
neg.QC = neg.QC[which(neg.QC$ID!="80025"),]
pos.QC = pos.QC[which(pos.QC$ID!="80025"),]

```
now, we have 14 pairs for analyses.


##Standardize the data according to placebos
```{r}
hist(neg.QC[,2])
#log-transform all the data
sample.QC$case <- factor(sample.QC$intervention,levels=c(1,2))
levels(sample.QC$case) <- c(0,1)
table(sample.QC$case,sample.QC$intervention)

info = sample.QC[,-c(4:5)]
marker = neg.QC[,1]

log.neg = log(neg.QC[,2:1530])
neg.lt = cbind(info,marker,log.neg)

log.pos = log(pos.QC[,2:1516])
pos.lt = cbind(info,marker,log.pos)

save(demographics, sample.QC, info, marker, neg.lt, pos.lt, file="Glance_logtrans.RData")
```

```{r}
info.marker = neg.lt[,1:5]

#normalization
table(sample.QC$case)

neg.norm <- apply(neg.lt[,6:1531],2,function(x) (x-mean(x[sample.QC$case==0],na.rm=T))/sd(x[sample.QC$case==0],na.rm=T))
mean(neg.norm[,1][sample.QC$case==0])
mean(neg.norm[,1][sample.QC$case==1])
neg.afternorm = cbind(info.marker,neg.norm)

pos.norm <- apply(pos.lt[,6:1517],2,function(x) (x-mean(x[sample.QC$case==0],na.rm=T))/sd(x[sample.QC$case==0],na.rm=T))
mean(pos.norm[,1][sample.QC$case==0])
mean(pos.norm[,1][sample.QC$case==1])
pos.afternorm = cbind(info.marker,pos.norm)

#Save RData
save(demographics, sample.QC, info.marker, neg.afternorm, pos.afternorm, file="Glance_afternorm.RData")
```

##check batch effect
```{r}
rm(list=ls())

load("Glance_afternorm.RData")

batch.neg<-neg.afternorm

library(ggplot2)
library(reshape2)
library(gridExtra)
library(data.table)


demo.select = demographics[which(demographics$ID!="80013"),]
demo.select = demo.select[which(demo.select$ID!="80025"),]
batch.neg.merge = merge(batch.neg, demo.select, by.x="id", by.y="ID", all=T)
dim(batch.neg.merge)
batch.neg.merge = batch.neg.merge[1:28,]
batch.neg.merge = batch.neg.merge[,-1540]
batch.neg.merge = batch.neg.merge[,-1532]
head(batch.neg.merge[,c("id","sex","order","case","age")])

######the plotting didn't work. Move on to PCA anyway.

plotdata.neg<-melt(batch.neg.merge, id=c("id","case","marker","sex","intervention","poisiton","order","age", "weight","height","bmi","bodyfat"))
p1 <- ggplot(plotdata.neg[plotdata.neg$sex==1,], aes(y=value,x=variable,fill=value)) + geom_tile() + scale_fill_gradient(low="lightblue", high="darkblue")+ xlab("metabolites") + ylab("obs")+theme(axis.ticks = element_blank(), axis.text.y = element_blank())+ggtitle("genes expression for male")
p0 <- ggplot(plotdata.neg[plotdata.neg$sex==0,], aes(y=value,x=variable,fill=value)) + geom_tile() + scale_fill_gradient(low="lightblue", high="darkblue")+ xlab("metabolites") + ylab("obs")+theme(axis.ticks = element_blank(), axis.text.y = element_blank())+ggtitle("genes expression for female")
grid.arrange( p0, p1, ncol=2)

p1 = ggplot(plotdata.neg[plotdata.neg$sex==1,], aes(y=plotdata.neg$value, x=plotdata.neg$variable)) + geom_tile() +ggtitle("genes expression for male")

```


#PCA to check batch effects
```{r}
library(stats)

library(ggplot2)
library(reshape2)
library(gridExtra)

#by sex and case
neg.pr.out = prcomp (neg.afternorm[,6:1531], scale =TRUE)
neg.pr.out$rotation[1:5,1:5]
pca.scores.neg = neg.pr.out$x
rownames(pca.scores.neg) <- neg.afternorm$marker
info.all = batch.neg.merge[,c("id","case","marker","sex","intervention","poisiton","order","age", "weight","height","bmi","bodyfat")]
pca.neg.mat = cbind(info.all,pca.scores.neg)

pca.plot.neg = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=as.factor(sex), shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites") + theme(panel.background = element_blank())

pos.pr.out = prcomp(pos.afternorm[,6:1517], scale=TRUE)
pos.pr.out$rotation[1:5,1:5]
pca.scores.pos = pos.pr.out$x
rownames(pca.scores.pos) = pos.afternorm$marker
pca.pos.mat = cbind(info.all, pca.scores.pos)

pca.plot.pos = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=as.factor(sex), shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg, pca.plot.pos, ncol=2)

#by order
pca.plot.neg2 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=as.factor(order), shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by order") + theme(panel.background = element_blank())

pca.plot.pos2 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=as.factor(order), shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by order") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg2, pca.plot.pos2, ncol=2)

#by bmi
info.all$bmi_c = NULL
info.all$bmi_c = ifelse(info.all$bmi>29.9, 1, 0)
table(info.all$bmi_c)

pca.neg.mat$bmi_c = info.all$bmi_c
pca.pos.mat$bmi_c = info.all$bmi_c

pca.plot.neg3 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=as.factor(bmi_c), shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by order") + theme(panel.background = element_blank())

pca.plot.pos3 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=as.factor(bmi_c), shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by order") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg3, pca.plot.pos3, ncol=2)

#save RData
save(demographics, sample.QC, info.marker, neg.afternorm, pos.afternorm, info.all, pca.pos.mat, pca.neg.mat, file="Glance_afternorm_batch.RData")

```

##Fold change
```{r}
rm(list=ls())

setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")
load("Glance_afternorm_batch.RData")

## May not need to do fold change because limma provide log2 FC as part of its output.

FoldChange.neg <- function(neg.afternorm, paired=FALSE, plot.hist=TRUE, 
    saveoutput=FALSE, outputname="fc.results") 
{
    # Read in data, collect groups information
    neg.afternorm <- editcolnames(neg.afternorm)
    groups <- factor(inputdata[, 1], levels=unique(inputdata[, 1]))
    grp_levs <- levels(groups)
    
    if (length(grp_levs) > 2) {
        stop(
            paste("The number of groups is greater than 2. Use", 
                " LinearModelFit() instead."
            )
        )
    }
    # Prepare empty matrices and populate
    folds <- matrix(NA, nrow=length(grp_levs), 
        ncol=ncol(inputdata) - 1, 
        dimnames=list(grp_levs, colnames(inputdata)[2:ncol(inputdata)])
    )
    grp_len <- c()
    for (ii in 1:length(grp_levs)) {
        grp_len <- c(grp_len, length(which(groups == levels(groups)[ii])))
    }
    
    new_mats <- c()
    for (ii in 1:length(grp_levs)) {
        new_mats[ii] <- list(inputdata[which(groups == levels(groups)[ii]), ])
    }
    
    # Perform fold change calculations
    if (!paired) {
        submeans <- c()
        means <- matrix(nrow=length(grp_levs), 
            ncol=length(colnames(inputdata[, -1])), 
            dimnames=list(grp_levs, colnames(inputdata[, -1]))
        )
        for (ii in 1:length(new_mats)) {
            submeans[ii] <- list(
                apply(new_mats[[ii]][, -1], 2, mean, na.rm=TRUE)
            )
            means[ii, ] <- submeans[[ii]]
        }
        for (ii in 1:length(means[, 1])) {
            for (jj in 1:length(means[1, ])) {
                folds[ii, jj] <- means[ii, jj] - means[1, jj]
            }
        }
    } else {
        folds[1, ] <- 1
        folds[2, ] <- apply(
            (new_mats[[2]][, -1] - new_mats[[1]][, -1]), 2, mean, na.rm=TRUE
        )
    }
    
    # Plot and/or save
    if (plot.hist) {
        hist(folds[2, ], breaks=50, xlab="Fold change", main="Histogram")
    }
    if (saveoutput) {
        write.csv(t(folds), paste(c(outputname, ".csv"), collapse=""))
    }
    
    return(t(folds))
}
```




## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

