---
title: "Metabolomics"
author: "wendyxwang"
date: "Sunday, May 24, 2015"
output:
  ioslides_presentation:
    fig_caption: yes
    fig_retina: 1
    keep_md: yes
    smaller: yes
  beamer_presentation: default
---

```{r,cache=FALSE}
# Set some global knitr options
library("knitr")
opts_chunk$set(tidy=TRUE, size='tiny',tidy.opts=list(blank=FALSE, width.cutoff=80), cache=TRUE, messages=FALSE)
```

## import all the data
```{r}
setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")

#read in subject demographics
library(xlsx)
demographics = read.xlsx("Subjects.xlsx",1)
colnames(demographics) = c("ID","baseline","order","age","sex","weight","height","bmi","bodyfat","notes")
summary(demographics)

#read in sample position and glucose
sample = read.csv("Glance_global_key_samples.csv", header=T, sep=",")
colnames(sample)
sample = sample[,-4]

#read in negative results
neg = read.csv("Neg.csv", header=T, sep=",", na.strings=".")
dim(neg)
sum(is.na(neg))
#impute all NA's with 1
neg[is.na(neg)] = 1
sum(neg == 1)

#read in positive results
pos = read.csv("Pos.csv", header=T, sep=",", na.strings=".")
dim(pos)
sum(is.na(pos))
pos[is.na(pos)] = 1
sum(pos == 1)

#Save RData
save.image(file="Glance.RData")
```


##Combine the dataset
```{r}
colnames(sample)
sample = sample[,1:5]
sample.QC = sample[!is.na(sample[,1]),]
dim(sample.QC)
neg$ID = sample.QC$id
pos$ID = sample.QC$id

sample.QC = sample.QC[which(substr(sample.QC$id,1,2)!="QC"),]
neg.QC = neg[which(substr(sample.QC$id,1,2)!="QC"),]
pos.QC = pos[which(substr(sample.QC$id,1,2)!="QC"),]

#exclude ill person
sample.QC = sample.QC[which(sample.QC$id!="80013"),]
neg.QC = neg.QC[which(neg.QC$ID!="80013"),]
pos.QC = pos.QC[which(pos.QC$ID!="80013"),]

#exclude drop-out
sample.QC = sample.QC[which(sample.QC$id!="80025"),]
neg.QC = neg.QC[which(neg.QC$ID!="80025"),]
pos.QC = pos.QC[which(pos.QC$ID!="80025"),]

```
now, we have 14 pairs for analyses.


##Standardize the data according to placebos
```{r}
hist(neg.QC[,2])
#log-transform all the data
sample.QC$case <- factor(sample.QC$intervention,levels=c(1,2))
levels(sample.QC$case) <- c(0,1)
table(sample.QC$case,sample.QC$intervention)

info = sample.QC[,-c(4:5)]
marker = neg.QC[,1]

log.neg = log(neg.QC[,2:1530])
neg.lt = cbind(info,marker,log.neg)

log.pos = log(pos.QC[,2:1516])
pos.lt = cbind(info,marker,log.pos)

save(demographics, sample.QC, info, marker, neg.lt, pos.lt, file="Glance_logtrans.RData")
```

```{r}
rm(list=ls())

load("Glance_logtrans.RData")

info.marker = neg.lt[,1:5]

#normalization
table(sample.QC$case)

hist(neg.lt[,6][sample.QC$case==0])
hist(neg.lt[,6][sample.QC$case==1])

neg.norm <- apply(neg.lt[,6:1531],2,function(x) (x-mean(x[sample.QC$case==0],na.rm=T))/sd(x[sample.QC$case==0],na.rm=T))
mean(neg.norm[,1][sample.QC$case==0])
mean(neg.norm[,1][sample.QC$case==1])
neg.afternorm = cbind(info.marker,neg.norm)

#code check: try a different method and get the same result
#neg.lt.t = t(neg.lt[,-c(1:5)])
#neg.norm.t = apply(neg.lt.t, 1, function(x) (x-mean(x[sample.QC$case==0]))/sd(x[sample.QC$case==00]))


pos.norm <- apply(pos.lt[,6:1517],2,function(x) (x-mean(x[sample.QC$case==0],na.rm=T))/sd(x[sample.QC$case==0],na.rm=T))
mean(pos.norm[,1][sample.QC$case==0])
mean(pos.norm[,1][sample.QC$case==1])
pos.afternorm = cbind(info.marker,pos.norm)

#Save RData
#decided to save non-normalized data for testing
save(demographics, sample.QC, info.marker, neg.lt, pos.lt, neg.afternorm, pos.afternorm, file="Glance_afternorm.RData")
```

##check batch effect
```{r}
rm(list=ls())

load("Glance_afternorm.RData")

batch.neg<-neg.afternorm

library(ggplot2)
library(reshape2)
library(gridExtra)
library(data.table)


demo.select = demographics[which(demographics$ID!="80013"),]
demo.select = demo.select[which(demo.select$ID!="80025"),]
batch.neg.merge = merge(batch.neg, demo.select, by.x="id", by.y="ID", all=T)
dim(batch.neg.merge)
batch.neg.merge = batch.neg.merge[1:28,]
batch.neg.merge = batch.neg.merge[,-1540]
batch.neg.merge = batch.neg.merge[,-1532]
head(batch.neg.merge[,c("id","sex","order","case","age")])

######the following plotting didn't work. Move on to PCA anyway.

plotdata.neg<-melt(batch.neg.merge, id=c("id","case","marker","sex","intervention","poisiton","order","age", "weight","height","bmi","bodyfat"))
p1 <- ggplot(plotdata.neg[plotdata.neg$sex==1,], aes(y=value,x=variable,fill=value)) + geom_tile() + scale_fill_gradient(low="lightblue", high="darkblue")+ xlab("metabolites") + ylab("obs")+theme(axis.ticks = element_blank(), axis.text.y = element_blank())+ggtitle("genes expression for male")
p0 <- ggplot(plotdata.neg[plotdata.neg$sex==0,], aes(y=value,x=variable,fill=value)) + geom_tile() + scale_fill_gradient(low="lightblue", high="darkblue")+ xlab("metabolites") + ylab("obs")+theme(axis.ticks = element_blank(), axis.text.y = element_blank())+ggtitle("genes expression for female")
grid.arrange( p0, p1, ncol=2)

p1 = ggplot(plotdata.neg[plotdata.neg$sex==1,], aes(y=plotdata.neg$value, x=plotdata.neg$variable)) + geom_tile() +ggtitle("genes expression for male")

```


#PCA to check batch effects
```{r}
library(stats)

library(ggplot2)
library(reshape2)
library(gridExtra)

#by sex and case
neg.pr.out = prcomp (neg.afternorm[,6:1531], scale =TRUE)
neg.pr.out$rotation[1:5,1:5]
pca.scores.neg = neg.pr.out$x
rownames(pca.scores.neg) <- neg.afternorm$marker
info.all = batch.neg.merge[,c("id","case","marker","sex","intervention","poisiton","order","age", "weight","height","bmi","bodyfat")]
info.all$sex = factor(info.all$sex, levels = c(0,1), labels = c("female", "male"))
info.all$order = factor(info.all$order, levels = c(1,2), labels = c("AB", "BA"))
pca.neg.mat = cbind(info.all,pca.scores.neg)

table(info.all$sex)
pca.neg.mat$sex = info.all$sex


pca.plot.neg = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=sex, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by sex") + theme(panel.background = element_blank())

pos.pr.out = prcomp(pos.afternorm[,6:1517], scale=TRUE)
pos.pr.out$rotation[1:5,1:5]
pca.scores.pos = pos.pr.out$x
rownames(pca.scores.pos) = pos.afternorm$marker
pca.pos.mat = cbind(info.all, pca.scores.pos)
pca.pos.mat$sex = info.all$sex

pca.plot.pos = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=sex, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by sex") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg, pca.plot.pos, ncol=2)

#by order
table(info.all$order)

pca.neg.mat$order = info.all$order
pca.pos.mat$order = info.all$order

pca.plot.neg2 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=order, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by order") + theme(panel.background = element_blank())

pca.plot.pos2 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=order, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by order") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg2, pca.plot.pos2, ncol=2)

#by bmi
info.all$bmi_c = NULL
info.all$bmi_c = ifelse(info.all$bmi>29.9, 1, 0)
info.all$bmi_c = factor(info.all$bmi_c, levels = c(0,1), labels = c("overweight", "obese"))
table(info.all$bmi_c)

pca.neg.mat$bmi_c = info.all$bmi_c
pca.pos.mat$bmi_c = info.all$bmi_c

pca.plot.neg3 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=bmi_c, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by BMI") + theme(panel.background = element_blank())

pca.plot.pos3 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=bmi_c, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by BMI") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg3, pca.plot.pos3, ncol=2)

#save RData
save(demographics, sample.QC, info.marker, neg.lt, pos.lt, neg.afternorm, pos.afternorm, info.all, pca.pos.mat, pca.neg.mat, file="Glance_afternorm_batch.RData")

```


##PCA (for visualization?)
```{r}
neg.pr.out = prcomp (neg.afternorm[,6:1531], scale =TRUE)
neg.pr.out$sdev[1:5]
neg.pr.var = neg.pr.out$sdev^2
neg.pve = neg.pr.var/sum(neg.pr.var)

pos.pr.out = prcomp(pos.afternorm[,6:1517], scale=TRUE)
pos.pr.out$sdev[1:5]
pos.pr.var = pos.pr.out$sdev^2
pos.pve = pos.pr.var/sum(pos.pr.var)

par(mfrow=c(1,2))
pve.plot.neg = plot(neg.pve, xlab=" Principal Component ", ylab=" Proportion of Variance Explained", ylim=c(0,1) ,type="b", main="negative metabolites")
pve.plot.pos = plot(pos.pve, xlab=" Principal Component ", ylab=" Proportion of Variance Explained", ylim=c(0,1) ,type="b", main="positive metabolites")

#cumulative pve
pve.plot.neg2 = plot(cumsum(neg.pve), xlab=" Principal Component ", ylab =" Cumulative Proportion of Variance Explained ", ylim=c(0,1) , type="b", main="negative metabolites")
pve.plot.pos2 = plot(cumsum(pos.pve), xlab=" Principal Component ", ylab =" Cumulative Proportion of Variance Explained ", ylim=c(0,1) , type="b", main="positive metabolites")

#no "elbows" for selecting PC's
```

##Fold change (not yet run)
```{r}
rm(list=ls())

setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")
load("Glance_afternorm_batch.RData")

## May not need to do fold change because limma provide log2 FC as part of its output.

FoldChange.neg <- function(neg.afternorm, paired=FALSE, plot.hist=TRUE, 
    saveoutput=FALSE, outputname="fc.results") 
{
    # Read in data, collect groups information
    neg.afternorm <- editcolnames(neg.afternorm)
    groups <- factor(inputdata[, 1], levels=unique(inputdata[, 1]))
    grp_levs <- levels(groups)
    
    if (length(grp_levs) > 2) {
        stop(
            paste("The number of groups is greater than 2. Use", 
                " LinearModelFit() instead."
            )
        )
    }
    # Prepare empty matrices and populate
    folds <- matrix(NA, nrow=length(grp_levs), 
        ncol=ncol(inputdata) - 1, 
        dimnames=list(grp_levs, colnames(inputdata)[2:ncol(inputdata)])
    )
    grp_len <- c()
    for (ii in 1:length(grp_levs)) {
        grp_len <- c(grp_len, length(which(groups == levels(groups)[ii])))
    }
    
    new_mats <- c()
    for (ii in 1:length(grp_levs)) {
        new_mats[ii] <- list(inputdata[which(groups == levels(groups)[ii]), ])
    }
    
    # Perform fold change calculations
    if (!paired) {
        submeans <- c()
        means <- matrix(nrow=length(grp_levs), 
            ncol=length(colnames(inputdata[, -1])), 
            dimnames=list(grp_levs, colnames(inputdata[, -1]))
        )
        for (ii in 1:length(new_mats)) {
            submeans[ii] <- list(
                apply(new_mats[[ii]][, -1], 2, mean, na.rm=TRUE)
            )
            means[ii, ] <- submeans[[ii]]
        }
        for (ii in 1:length(means[, 1])) {
            for (jj in 1:length(means[1, ])) {
                folds[ii, jj] <- means[ii, jj] - means[1, jj]
            }
        }
    } else {
        folds[1, ] <- 1
        folds[2, ] <- apply(
            (new_mats[[2]][, -1] - new_mats[[1]][, -1]), 2, mean, na.rm=TRUE
        )
    }
    
    # Plot and/or save
    if (plot.hist) {
        hist(folds[2, ], breaks=50, xlab="Fold change", main="Histogram")
    }
    if (saveoutput) {
        write.csv(t(folds), paste(c(outputname, ".csv"), collapse=""))
    }
    
    return(t(folds))
}
```



## adjut for gender, age and bmi
```{r}
rm(list=ls())

setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")
load("Glance_afternorm_batch.RData")


library(reshape)
sub.ma <- neg.afternorm[,5:1531]
sub.ma.t = t(sub.ma)
colnames(sub.ma.t) = sub.ma.t[1,]
sub.ma.t = sub.ma.t[-1,]

sub.info <- info.all
info.temp <- sub.info
neg.compound.name = colnames(neg.afternorm)[6:1531]

info.temp$sex <- as.factor(info.temp$sex)
info.temp$case <- as.factor(info.temp$case)
info.temp$bmi_c <- as.factor(info.temp$bmi_c)
info.temp$age <- as.factor(info.temp$age)

#require library
library(qvalue)

estim.casecont <- function(mylm,info.temp,m){
  coefs <- summary(mylm)$coef[,1]
  name.coefs <- names(coefs)
   # gender #
  sex.name <- name.coefs[1:length(name.coefs) %in% grep("sex",name.coefs)]
  sex.name <- unlist(lapply(strsplit(sex.name,"sex"),function(x) x[2]))
  k <- 1
  X <- ifelse(as.character(info.temp$sex)==sex.name[k],1,0)
  if(length(sex.name)>1){
    for(k in 2:length(sex.name))  X <- cbind(X,ifelse(as.character(info.temp$sex)==sex.name[k],1,0))
    table(X[,ncol(X)-1],info.temp$sex)
    table(X[,ncol(X)],info.temp$sex)
    }
  # bmi_c #
  bmi.name <- name.coefs[1:length(name.coefs) %in% grep("bmi",name.coefs)]
  age.name <- unlist(lapply(strsplit(bmi.name,"bmi"),function(x) x[2]))
  k <- 1
  X <- cbind(X,ifelse(as.character(info.temp$bmi_c)==bmi.name[k],1,0))
  table(X[,ncol(X)],info.temp$bmi_c)
  # age #
  age.name <- name.coefs[1:length(name.coefs) %in% grep("age",name.coefs)]
  age.name <- unlist(lapply(strsplit(age.name,"age"),function(x) x[2]))
  k <- 1
  X <- cbind(X,ifelse(as.character(info.temp$age)==age.name[k],1,0))
  for(k in 2:length(age.name))  X <- cbind(X,ifelse(as.character(info.temp$age)==age.name[k],1,0))
  table(X[,ncol(X)-1],info.temp$age)
  table(X[,ncol(X)],info.temp$age)


  est.m <- m - coefs[-c(1,2)]%*%t(X)
  return(est.m)
}


lm.fun <- function(sub.ma.t,sub.info,neg.compound.name){
  sub.info$sex <- as.factor(sub.info$sex)
  sub.info$case <- as.factor(sub.info$case)
  sub.info$bmi_c <- as.factor(sub.info$bmi_c)
  sub.info$age <- as.factor(sub.info$age)
  
  ## linear regression to M with a
  coef <- pvalue <- NA
  est.m<-matrix(rep(NA,nrow(sub.ma.t)*ncol(sub.ma.t)),nrow=nrow(sub.ma.t))
  i <- 1
  for(i in 1:nrow(sub.ma.t)){
    m.temp <- as.numeric(sub.ma.t[i,])

    mylm <- lm(m.temp ~ case+sex+bmi_c+age, data=sub.info)
    coef[i] <- mylm$coef["case1"]
    pvalue[i] <- summary(mylm)$coef["case1",4]
    est.m[i,] <- estim.casecont(mylm,sub.info,m.temp)    
  }
  p.d <- data.frame(ind=1:length(pvalue),pvalue)
  q.d <- data.frame(ind=p.d$ind[!is.na(p.d$p)],q=qvalue(na.omit(p.d$p),robust=T)$qvalues)
  qvalue <- merge(p.d,q.d,by="ind",all.x=T,all.y=T)$q

  result.all <- data.frame(neg.compound.name=neg.compound.name,coef=coef,pvalue=pvalue,qvalue=qvalue)
  result.all <- result.all[order(result.all$pvalue),]
#  return(result.all)
  return(list(result=result.all,est=est.m))  
}

result<-lm.fun(sub.ma.t,sub.info,neg.compound.name)
#ls(result)
neg.est.m<-result[["est"]]
neg.est.result = result[["result"]]
```

## do the same thing for positive data
```{r}
sub.ma <- pos.afternorm[,5:1517]
sub.ma.t = t(sub.ma)
colnames(sub.ma.t) = sub.ma.t[1,]
sub.ma.t = sub.ma.t[-1,]

sub.info <- info.all
info.temp <- sub.info
pos.compound.name = colnames(pos.afternorm)[6:1517]

info.temp$sex <- as.factor(info.temp$sex)
info.temp$case <- as.factor(info.temp$case)
info.temp$bmi_c <- as.factor(info.temp$bmi_c)
info.temp$age <- as.factor(info.temp$age)

# estim.casecont and lm.fun is exactly the same function

result<-lm.fun(sub.ma.t,sub.info,pos.compound.name)
#ls(result)
pos.est.m<-result[["est"]]
pos.est.result = result[["result"]]

#save RData
save(demographics, sample.QC, info.marker, neg.lt, pos.lt, neg.afternorm, pos.afternorm, info.all, pca.pos.mat, pca.neg.mat, neg.est.m, neg.est.result, pos.est.m, pos.est.result, file="Glance_afternorm_batch_lm.RData")

```

##Moderated t-test
```{r}
rownames(neg.est.m) <- colnames(neg.afternorm)[6:1531]
colnames(neg.est.m) <- neg.afternorm[,5]

rownames(pos.est.m) <- colnames(pos.afternorm)[6:1517]
colnames(pos.est.m) <- pos.afternorm[,5]

library(limma)
#negative data

est.m = neg.est.m
mm <-model.matrix(~0+id+case, data=info.temp)
mm <- mm[,-7]
mm <- mm[, -(16:20)]
mm <- mm[,-13]
fit <- lmFit(est.m, mm)

ebay <- eBayes(fit)
top <- topTable(ebay, coef="case1", number=Inf, sort.by="p")
head(top)
mean(top$adj.P.Val<0.01)
summary(top$adj.P.Val)
summary(top$P.Value)
hist(top$P.Value, freq=F, main="Histogram of unadjusted p-values", xlab='p-values', breaks=100)
abline(v=0.05, col=2, lwd=2, lty=2)

neg.top = top

#try neg.lt
library(reshape)
sub.ma <- neg.lt[,6:1531]
sub.ma.t = t(sub.ma)
colnames(sub.ma.t) = sub.ma.t[1,]
sub.ma.t = sub.ma.t[-1,]

sub.info <- info.all
info.temp <- sub.info
neg.compound.name = colnames(neg.lt)[6:1534]

info.temp$sex <- as.factor(info.temp$sex)
info.temp$case <- as.factor(info.temp$case)
info.temp$bmi_c <- as.factor(info.temp$bmi_c)
info.temp$age <- as.factor(info.temp$age)

est.m = sub.ma.t
mm <-model.matrix(~0+id+case, data=info.temp)
mm <- mm[,-7]
mm <- mm[, -(16:20)]
mm <- mm[,-13]
fit <- lmFit(est.m, mm)

ebay <- eBayes(fit)
top <- topTable(ebay, coef="case1", number=Inf, sort.by="p")
head(top)
mean(top$adj.P.Val<0.01)
summary(top$adj.P.Val)
summary(top$P.Value)
hist(top$P.Value, freq=F, main="Histogram of unadjusted p-values", xlab='p-values', breaks=100)
abline(v=0.05, col=2, lwd=2, lty=2)
#it's the same
```

```{r}
#positive data
est.m = pos.est.m

fit <- lmFit(est.m, mm)

ebay <- eBayes(fit)
top <- topTable(ebay, coef="case1", number=Inf, sort.by="p")
head(top)
mean(top$adj.P.Val<0.01)
summary(top$adj.P.Val)
summary(top$P.Value)
hist(top$P.Value, freq=F, main="Histogram of unadjusted p-values", xlab='p-values', breaks=100)
abline(v=0.05, col=2, lwd=2, lty=2)

hist(top$adj.P.Val, freq=F)

pos.top = top


save(demographics, sample.QC, info.marker, neg.lt, pos.lt, neg.afternorm, pos.afternorm, info.all, pca.pos.mat, pca.neg.mat, neg.est.m, neg.est.result, pos.est.m, pos.est.result, neg.top, pos.top, file="Glance_afternorm_batch_lm_top.RData")
```
nothing significant at all!

##try regular t-test
```{r}
#negative
rownames(neg.est.m) <- colnames(neg.afternorm)[6:1531]
colnames(neg.est.m) <- neg.afternorm[,5]

```

```{r paired t testing,echo=FALSE,cache=TRUE}

table(colnames(neg.est.m)==info.all$marker)
info.all$id <- as.character(info.all$id)
info.all[1,]
table(table(info.all$id))
case.m <- neg.est.m[,info.all$case==1]
ctrl.m <- neg.est.m[,info.all$case==0]

case.info <- info.all[info.all$case==1,]
ctrl.info <- info.all[info.all$case==0,]

table(case.info$id==ctrl.info$id)

p.t <- sapply(1:nrow(case.m),function(x) t.test(case.m[x,],ctrl.m[x,],paired=T)$p.value)

effect.size <- sapply(1:nrow(case.m),function(x) (mean(case.m[x,],na.rm=T)-mean(ctrl.m[x,],na.rm=T))/sd(ctrl.m[x,],na.rm=T))
qvalue <- qvalue(p.t,robust=T)$qvalues
p.t.bh<-p.adjust(p.t, method = "BH")

plot(effect.size,p.t,pch=19,col="blue",xlab="Effect Size",ylab="p value paired T",main=paste(sum(info.all$case==1)," GC treated vs. ",sum(info.all$case==0)," Placebo",sep=""))

plot(effect.size,p.t.bh,pch=19,col="blue",xlab="Effect Size",ylab="p value paired T",main=paste(sum(info.all$case==1)," GC treated vs. ",sum(info.all$case==0)," Placebo",sep=""))

plot(effect.size,qvalue,pch=19,col="blue",xlab="Effect Size",ylab="p value paired T",main=paste(sum(info.all$case==1)," GC treated vs. ",sum(info.all$case==0)," Placebo",sep=""))

plot(result$result$coef,result$result$pvalue,pch=19,col="blue",xlab="coef",ylab="p",main=paste(sum(info.all$case==1)," GC treated vs. ",sum(info.all$case==0)," Placebo",sep=""))
```


##pathway analysis
```{r}
rm(list=ls())
load("Glance_afternorm_batch_lm_top.RData")

neg.compound.name = colnames(neg.afternorm)[6:1531]
neg.top.name = cbind(neg.compound.name, neg.top)

pos.compound.name = colnames(pos.afternorm)[6:1517]
pos.top.name = cbind(pos.compound.name, pos.top)

write.csv(neg.top.name, row.names=F, file="neg_top_name.csv")
write.csv(pos.top.name, file="pos_top_name.csv")

#MetaMapR
library(shiny)
shiny::runGitHub('MetaMapR','dgrapov')
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

