---
title: "Final Plan"
author: "wendyxwang"
date: "Tuesday, June 02, 2015"
output: html_document.
---

Data preprocessing and analyses have changed:
1. Log-transform the data
2. Check boxplots of 5 QC's -> check the array performance
3. Remove data with >50% missing
4. For the rest of missing, impute with 1/2 minimum value
5. Boxplot of the subjects.
6. Histogram of the metabolites (first 5 by case status)
7. Wilcoxon Rank Test: paired, tied, no correction for continuity, effec size
8. Exploratory:
8.1 PCA
8.2 Clustering

## import all the data
```{r}
rm(list=ls())

setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")

#read in subject demographics
library(xlsx)
demographics = read.xlsx("Subjects.xlsx",1)
colnames(demographics) = c("ID","baseline","order","age","sex","weight","height","bmi","bodyfat","notes")
summary(demographics)

#read in sample position and glucose
sample = read.csv("Glance_global_key_samples.csv", header=T, sep=",")
colnames(sample)
sample = sample[,1:3]

#read in negative results
neg = read.csv("Neg.csv", header=T, sep=",", na.strings=".", check.names=FALSE)
dim(neg)
#36 1530
sum(is.na(neg))
# 19995

#read in positive results
pos = read.csv("Pos.csv", header=T, sep=",", na.strings=".", check.names=FALSE)
dim(pos)
#36 1516
sum(is.na(pos))
#19123

#Save RData
save.image(file="Glance.RData")
```

##Exclude 80013 and 80025
```{r}
colnames(sample)

sample.QC = sample[!is.na(sample[,1]),]
dim(sample.QC)
ID = sample.QC$id

neg.QC = cbind(ID, neg)
pos.QC = cbind(ID, pos)

#exclude ill person
sample.QC = sample.QC[which(sample.QC$id!="80013"),]
neg.QC = neg.QC[which(neg.QC$ID!="80013"),]
pos.QC = pos.QC[which(pos.QC$ID!="80013"),]

#exclude drop-out
sample.QC = sample.QC[which(sample.QC$id!="80025"),]
neg.QC = neg.QC[which(neg.QC$ID!="80025"),]
pos.QC = pos.QC[which(pos.QC$ID!="80025"),]

table(sample.QC$id)
#now we have 5 QCs and 14 pairs

save(demographics, sample.QC, neg.QC, pos.QC, file="Glance_drop.RData")
```

# 1. Log-transform the data
```{r}
rm(list=ls())

load("Glance_drop.RData")

log.neg.QC = log(neg.QC[,3:1531])
log.pos.QC = log(pos.QC[,3:1517])

sum(is.na(log.neg.QC))
# 18616
sum(is.na(log.pos.QC))
#17663
```

# 2. Check boxplots of 5 QC's -> check the array performance
```{r}
rowname.QC = paste("QC", c(1:5), sep="")
#negative
QC = log.neg.QC[which(substr(sample.QC$id,1,2)=="QC"),]
rownames(QC) = rowname.QC
boxplot(t(QC))

#positive
QC = log.pos.QC[which(substr(sample.QC$id,1,2)=="QC"),]
rownames(QC) = rowname.QC
boxplot(t(QC))

save(demographics, sample.QC, neg.QC, pos.QC, log.neg.QC, log.pos.QC, file="Glance_log_QCcheck.RData")
```
Array behaved well.

#3. Remove data with >50% missing
```{r}
rm(list=ls())

load("Glance_log_QCcheck.RData")

#first remove QCs, they are done.
sample.QC = sample.QC[which(substr(sample.QC$id,1,2)!="QC"),]
log.neg.QC = log.neg.QC[which(substr(sample.QC$id,1,2)!="QC"),]
log.pos.QC = log.pos.QC[which(substr(sample.QC$id,1,2)!="QC"),]

ID.marker = neg.QC[,c(1:2)][which(substr(neg.QC$ID,1,2)!="QC"),]

#neg
filter.neg = apply(log.neg.QC, 2, function(x) sum(is.na(x)))
summary(filter.neg)
table(filter.neg)
#482 metabolites have >14 NA's
neg.f = log.neg.QC[,which(filter.neg<=14)]
#now have 1047

#pos
filter.pos = apply(log.pos.QC, 2, function(x) sum(is.na(x)))
summary(filter.pos)
table(filter.pos)
#523 metabolites have >14 NA's
pos.f = log.pos.QC[,which(filter.pos<=14)]
#now have 992

save(demographics, sample.QC, ID.marker, neg.f, pos.f, file="Glance_dropmissing.RData")
#also save prob.filter
save(demographics, sample.QC, ID.marker, neg.f, pos.f,prob.filter, file="Glance_dropmissing_probfilter.RData")
```

#4. For the rest of missing, impute with 1/2 minimum value
```{r}
rm(list=ls())

load("Glance_dropmissing.RData")

#neg
sum(is.na(neg.f))
#6921
neg.min = apply(neg.f, 2, function(x) min(x, na.rm=T))

neg.im = apply(neg.f,2,function(x) ifelse(is.na(x),min(x, na.rm=T)/2,x))
sum(is.na(neg.im))
neg.min.im = apply(neg.im, 2, function(x) min(x))

neg.min[1041:1047]
neg.min.im[1041:1047]

#pos
sum(is.na(pos.f))
#5180
pos.min = apply(pos.f, 2, function(x) min(x, na.rm=T))

pos.im = apply(pos.f,2,function(x) ifelse(is.na(x),min(x, na.rm=T)/2,x))
sum(is.na(pos.im))
neg.pos.im = apply(pos.im, 2, function(x) min(x))

pos.min[985:992]
neg.pos.im[985:992]

save(demographics, sample.QC, ID.marker, neg.im, pos.im, file="Glance_impute.RData")
```

#5. Boxplot of the subjects.
```{r}
rm(list=ls())

load("Glance_impute.RData")

sample.QC$case <- factor(sample.QC$intervention,levels=c(1,2))
levels(sample.QC$case) <- c(0,1)
table(sample.QC$case,sample.QC$intervention)

boxplot(t(neg.im))
boxplot(t(pos.im))

par(mfrow=c(2,2))
boxplot(t(neg.im[which(sample.QC$case==1),]),main="Negative G&C")
boxplot(t(pos.im[which(sample.QC$case==1),]),main="Positive G&C")
boxplot(t(neg.im[which(sample.QC$case==0),]),main="Negative Placebo")
boxplot(t(pos.im[which(sample.QC$case==0),]),main="Positive Placebo")

dev.off()


```

#6. Histogram of the metabolites (first 5)
```{r}
par(mfrow=c(2,5))
hist(neg.im[,1], main="M1", xlab="negative")
hist(neg.im[,2], main="M2", xlab="negative")
hist(neg.im[,3], main="M3", xlab="negative")
hist(neg.im[,4], main="M4", xlab="negative")
hist(neg.im[,5], main="M5", xlab="negative")

hist(pos.im[,1], main="M1", xlab="positive")
hist(pos.im[,2], main="M2", xlab="positive")
hist(pos.im[,3], main="M3", xlab="positive")
hist(pos.im[,4], main="M4", xlab="positive")
hist(pos.im[,5], main="M5", xlab="positive")

#by case
#neg
par(mfrow=c(2,5))
hist(neg.im[,1][sample.QC$case==1], main="M1", xlab="G&C")
hist(neg.im[,2][sample.QC$case==1], main="M2", xlab="G&C")
hist(neg.im[,3][sample.QC$case==1], main="M3", xlab="G&C")
hist(neg.im[,4][sample.QC$case==1], main="M4", xlab="G&C")
hist(neg.im[,5][sample.QC$case==1], main="M5", xlab="G&C")

hist(neg.im[,1][sample.QC$case==0], main="M1", xlab="Placebo")
hist(neg.im[,2][sample.QC$case==0], main="M2", xlab="Placebo")
hist(neg.im[,3][sample.QC$case==0], main="M3", xlab="Placebo")
hist(neg.im[,4][sample.QC$case==0], main="M4", xlab="Placebo")
hist(neg.im[,5][sample.QC$case==0], main="M5", xlab="Placebo")

#pos
par(mfrow=c(2,5))
hist(pos.im[,1][sample.QC$case==1], main="M1", xlab="G&C")
hist(pos.im[,2][sample.QC$case==1], main="M2", xlab="G&C")
hist(pos.im[,3][sample.QC$case==1], main="M3", xlab="G&C")
hist(pos.im[,4][sample.QC$case==1], main="M4", xlab="G&C")
hist(pos.im[,5][sample.QC$case==1], main="M5", xlab="G&C")

hist(pos.im[,1][sample.QC$case==0], main="M1", xlab="Placebo")
hist(pos.im[,2][sample.QC$case==0], main="M2", xlab="Placebo")
hist(pos.im[,3][sample.QC$case==0], main="M3", xlab="Placebo")
hist(pos.im[,4][sample.QC$case==0], main="M4", xlab="Placebo")
hist(pos.im[,5][sample.QC$case==0], main="M5", xlab="Placebo")

dev.off()

save(demographics, sample.QC, ID.marker, neg.im, pos.im, file="Glance_ready.RData")
```

#7. Wilcoxon Rank Test: paired, tied, no correction for continuity, effec size
```{r}
rm(list=ls())

load("Glance_ready.RData")

library(coin)

case = sample.QC$case

#neg
neg.wilcox = apply(neg.im, 2, function(x) wilcoxsign_test(x[which(case==1)] ~ x[which(case==0)], zero.method="Wilcoxon", distribution="asympt"))
neg.p = unlist(lapply(neg.wilcox, function(x) pvalue(x)))
neg.es = unlist(lapply(neg.wilcox, function(x) statistic(x)))

list(neg.wilcox[[1]])

hist(neg.p)
hist(neg.es)

#pos
pos.wilcox = apply(pos.im, 2, function(x) wilcoxsign_test(x[which(case==1)] ~ x[which(case==0)], zero.method="Wilcoxon", distribution="asympt"))
pos.p = unlist(lapply(pos.wilcox, function(x) pvalue(x)))
pos.es = unlist(lapply(pos.wilcox, function(x) statistic(x)))

list(pos.wilcox[[1]])

hist(pos.p)
hist(pos.es)

par(mfrow=c(2,2))
hist(neg.p, freq=F, main="Unadjusted p-values (Negative)", xlab='p-values', breaks=50)
abline(v=0.05, col=2, lwd=2, lty=2)
hist(neg.es, freq=F, main="Test statistics (Negative)", xlab="z-scores", breaks=50)
hist(pos.p, freq=F, main="Unadjusted p-values (Positive)", xlab='p-values', breaks=50)
abline(v=0.05, col=2, lwd=2, lty=2)
hist(pos.es, freq=F, main="Test statistics (Positive)", xlab="z-scores", breaks=50)

dev.off()

#create result datasets
neg.result = cbind(neg.es, neg.p)
pos.result = cbind(pos.es, pos.p)

#clean the rownames
rownames(neg.result) = colnames(neg.im)
rownames(pos.result) = colnames(pos.im)


save(demographics, sample.QC, ID.marker, neg.im, pos.im, neg.wilcox, pos.wilcox, neg.result, pos.result, file="Glance_afterWilcox.RData")
```

```{r}

#unadjusted p<0.05
neg.unadj.p = neg.result[which(data.frame(neg.result)$neg.p<0.05),]
neg.unadj.p = neg.unadj.p[order(data.frame(neg.unadj.p)$neg.p),]
#27
pos.unadj.p = pos.result[which(data.frame(pos.result)$pos.p<0.05),]
pos.unadj.p = pos.unadj.p[order(data.frame(pos.unadj.p)$pos.p),]
#39

#fdr adjustment
neg.p.adjust = p.adjust(data.frame(neg.result)$neg.p, method ="BH")
summary(neg.p.adjust)

pos.p.adjust = p.adjust(data.frame(pos.result)$pos.p, method ="BH")
summary(pos.p.adjust)
#nothing significant

#what if we order p-value first
#neg.result = data.frame(neg.result)
#neg.result = neg.result[order(neg.result$neg.p),]
#neg.p.adj = p.adjust(neg.result$neg.p, method="BH")
#summary(neg.p.adj)
#exactly the same

write.csv(neg.unadj.p, row.names=T,file="Negative_unadjusted_pvalues.csv")
write.csv(pos.unadj.p, row.names=T,file="Positive_unadjusted_pvalues.csv")
```

```{r}
library(shiny)
shiny::runGitHub('MetaMapR','dgrapov')

```

## 8. Exploratory:
# 8.1 PCA
```{r}
#use imputed data
rm(list=ls())

load("Glance_afterWilcox.RData")


library(ggplot2)
library(reshape2)
library(gridExtra)
library(RColorBrewer)

neg.pr.out = prcomp (neg.im, scale =TRUE)
neg.pr.out$rotation[1:5,1:5]
pca.scores.neg = neg.pr.out$x

sample.QC$caseid = paste(sample.QC$case, sample.QC$id, sep=".")
rownames(pca.scores.neg) <- sample.QC$caseid

#by case
demo.select = demographics[which(demographics$ID!="80013"),]
demo.select = demo.select[which(demo.select$ID!="80025"),]
demo.select = demo.select[-c(1:4),]
info.temp = merge(sample.QC, demo.select, by.x="id", by.y="ID", all=T)
info.temp = info.temp[,-14]
info.all = merge(info.temp, ID.marker, by.x="id", by.y="ID", all=T)

pca.neg.mat = cbind(info.all,pca.scores.neg)
pca.neg.mat$case = as.factor(pca.neg.mat$case)
pca.neg.mat$sex = factor(pca.neg.mat$sex, levels = c(0,1), labels = c("female", "male"))
pca.neg.mat$order = factor(pca.neg.mat$order, levels = c(1,2), labels = c("AB", "BA"))

neg.pca1 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=case)) + geom_point(shape=16,size=3, aes(colour=case)) + ggtitle("PCA for negative metabolites by case") + theme(panel.background = element_blank())
neg.pca2 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC3, color=case)) + geom_point(shape=16,size=3) + ggtitle("PCA for negative metabolites by case") + theme(panel.background = element_blank())
neg.pca3 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC3, color=case)) + geom_point(shape=16,size=3) + ggtitle("PCA for negative metabolites by case") + theme(panel.background = element_blank())
neg.pca4 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC4, color=case)) + geom_point(shape=16,size=3) + ggtitle("PCA for negative metabolites by case") + theme(panel.background = element_blank())

grid.arrange(neg.pca1, neg.pca2, neg.pca3, neg.pca4, ncol=2)

#pos
pos.pr.out = prcomp (pos.im, scale =TRUE)
pos.pr.out$rotation[1:5,1:5]
pca.scores.pos = pos.pr.out$x

rownames(pca.scores.pos) <- sample.QC$caseid

pca.pos.mat = cbind(info.all,pca.scores.pos)
pca.pos.mat$case = as.factor(pca.pos.mat$case)
pca.pos.mat$sex = factor(pca.pos.mat$sex, levels = c(0,1), labels = c("female", "male"))
pca.pos.mat$order = factor(pca.pos.mat$order, levels = c(1,2), labels = c("AB", "BA"))

pos.pca1 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=case)) + geom_point(shape=16,size=3, aes(colour=case)) + ggtitle("PCA for positive metabolites by case") + theme(panel.background = element_blank())
pos.pca2 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC3, color=case)) + geom_point(shape=16,size=3) + ggtitle("PCA for positive metabolites by case") + theme(panel.background = element_blank())
pos.pca3 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC3, color=case)) + geom_point(shape=16,size=3) + ggtitle("PCA for positive metabolites by case") + theme(panel.background = element_blank())
pos.pca4 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC4, color=case)) + geom_point(shape=16,size=3) + ggtitle("PCA for positive metabolites by case") + theme(panel.background = element_blank())

grid.arrange(pos.pca1, pos.pca2, pos.pca3, pos.pca4, ncol=2)

#by sex and case

pca.plot.neg = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=sex, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by sex") + theme(panel.background = element_blank())

pos.pr.out = prcomp(pos.afternorm[,6:1517], scale=TRUE)
pos.pr.out$rotation[1:5,1:5]
pca.scores.pos = pos.pr.out$x
rownames(pca.scores.pos) = pos.afternorm$marker
pca.pos.mat = cbind(info.all, pca.scores.pos)
pca.pos.mat$sex = info.all$sex

pca.plot.pos = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=sex, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by sex") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg, pca.plot.pos, ncol=2)

#by order
table(info.all$order)

pca.neg.mat$order = info.all$order
pca.pos.mat$order = info.all$order

pca.plot.neg2 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=order, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by order") + theme(panel.background = element_blank())

pca.plot.pos2 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=order, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by order") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg2, pca.plot.pos2, ncol=2)

#by bmi
info.all$bmi_c = NULL
info.all$bmi_c = ifelse(info.all$bmi>29.9, 1, 0)
info.all$bmi_c = factor(info.all$bmi_c, levels = c(0,1), labels = c("overweight", "obese"))
table(info.all$bmi_c)

pca.neg.mat$bmi_c = info.all$bmi_c
pca.pos.mat$bmi_c = info.all$bmi_c

pca.plot.neg3 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=bmi_c, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by BMI") + theme(panel.background = element_blank())

pca.plot.pos3 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=bmi_c, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by BMI") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg3, pca.plot.pos3, ncol=2)

#save RData
save(demographics, sample.QC, ID.marker, neg.im, pos.im, neg.wilcox, pos.wilcox, neg.result, pos.result, info.all, pca.pos.mat, pca.neg.mat, file="Glance_afterWilcox_PCA.RData")

```
8.2 Clustering
(Ben)

Additional
#Calculate Fold Change
```{r}
rm(list=ls())

load("Glance_afterWilcox_PCA.RData")


library(ggplot2)
library(reshape2)
library(gridExtra)
library(RColorBrewer)

neg.fc = neg.im[which(sample.QC$case==1),] - neg.im[which(sample.QC$case==0),]
pos.fc = pos.im[which(sample.QC$case==1),] - pos.im[which(sample.QC$case==0),]

save(list=ls(), file="Glance_afterwilcox_fc.RData")

```

#Extract the ones with p<0.01
```{r}
setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")

load("Glance_afterwilcox_fc.RData")

neg.list = neg.result[which(neg.result[,2]<0.01),]
#3
pos.list = pos.result[which(pos.result[,2]<0.01),]
#4

info.temp = info.all[which(info.all$case==1),]
info.temp = info.temp[!duplicated(info.temp$id),]
#info.temp = info.temp[which(row.names(info.temp)==row.names(neg.fc)),]
bfp.med = median(info.temp$bodyfat)

#neg
neg.fc.list = neg.fc[,which(colnames(neg.fc)==rownames(neg.list))]
neg.up = apply(neg.fc.list, 2, function(x) sum(x>0))
neg.mean = apply(neg.fc.list, 2, function(x) mean(x))
neg.male = apply(neg.fc.list[info.temp$sex==1,], 2, function(x) mean(x))
neg.female = apply(neg.fc.list[info.temp$sex==0,], 2, function(x) mean(x))
neg.bfplow =apply(neg.fc.list[info.temp$bodyfat<bfp.med,],2, function(x) mean(x))
neg.bfphigh=apply(neg.fc.list[info.temp$bodyfat>=bfp.med,],2,function(x) mean(x))

neg.table = cbind(neg.list, neg.up, neg.mean, neg.male, neg.female, neg.bfplow, neg.bfphigh)

#pos
a = rownames(pos.list)
pos.fc.list = pos.fc[,which(colnames(pos.fc) %in% a)]
pos.up = apply(pos.fc.list, 2, function(x) sum(x>0))
pos.mean = apply(pos.fc.list, 2, function(x) mean(x))
pos.male = apply(pos.fc.list[info.temp$sex==1,], 2, function(x) mean(x))
pos.female = apply(pos.fc.list[info.temp$sex==0,], 2, function(x) mean(x))
pos.bfplow =apply(pos.fc.list[info.temp$bodyfat<bfp.med,],2, function(x) mean(x))
pos.bfphigh=apply(pos.fc.list[info.temp$bodyfat>=bfp.med,],2,function(x) mean(x))

pos.table = cbind(pos.list, pos.up, pos.mean, pos.male, pos.female, pos.bfplow, pos.bfphigh)

write.csv(neg.table, row.names=T,file="Negative_table.csv")
write.csv(pos.table, row.names=T,file="Positive_table.csv")

save(demographics, sample.QC, ID.marker, neg.im, pos.im, neg.wilcox, pos.wilcox, neg.result, pos.result, info.all, pca.pos.mat, pca.neg.mat,info.temp, neg.fc, pos.fc, neg.table, pos.table, file="Glance_afterWilcox_fc_table.RData")

```
#PCA on FC data
```{r}
setwd("E:/Course2015Spring/BIOST578/Project/FInal_Project_Metabolomics/Data")

load("Glance_afterwilcox_fc.RData")

library(ggplot2)
library(reshape2)
library(gridExtra)
library(RColorBrewer)

neg.pr.fc = prcomp (neg.fc, scale =TRUE)
neg.pr.fc$rotation[1:5,1:5]
pca.scores.neg.fc = neg.pr.fc$x

sample.QC$caseid = paste(sample.QC$case, sample.QC$id, sep=".")
rownames(pca.scores.neg.fc) <- sample.QC$id[which(sample.QC$case==1)]


demo.select = demographics[which(demographics$ID!="80013"),]
demo.select = demo.select[which(demo.select$ID!="80025"),]
demo.select = demo.select[-c(1:4),]
info.temp = merge(sample.QC, demo.select, by.x="id", by.y="ID", all=T)
info.temp = info.temp[,-14]
info.all = merge(info.temp, ID.marker, by.x="id", by.y="ID", all=T)

info.all.fc = info.all[which(sample.QC$case==1),]

pca.neg.mat.fc = cbind(info.all.fc,pca.scores.neg.fc)

pca.neg.mat.fc$sex = factor(pca.neg.mat.fc$sex, levels = c(0,1), labels = c("female", "male"))
pca.neg.mat.fc$order = factor(pca.neg.mat.fc$order, levels = c(1,2), labels = c("AB", "BA"))

#pos
pos.pr.fc = prcomp (pos.fc, scale =TRUE)
pos.pr.fc$rotation[1:5,1:5]
pca.scores.pos.fc = pos.pr.fc$x

rownames(pca.scores.pos.fc) <- sample.QC$id[which(sample.QC$case==1)]

pca.pos.mat.fc = cbind(info.all.fc,pca.scores.pos.fc)

pca.pos.mat.fc$sex = factor(pca.pos.mat.fc$sex, levels = c(0,1), labels = c("female", "male"))
pca.pos.mat.fc$order = factor(pca.pos.mat.fc$order, levels = c(1,2), labels = c("AB", "BA"))

#by sex

pca.plot.neg.fc1 = ggplot(data=pca.neg.mat.fc,aes(x=PC1, y=PC2, color=sex)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by sex") + theme(panel.background = element_blank())
pca.plot.neg.fc2 = ggplot(data=pca.neg.mat.fc,aes(x=PC1, y=PC3, color=sex)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by sex") + theme(panel.background = element_blank())
pca.plot.neg.fc3 = ggplot(data=pca.neg.mat.fc,aes(x=PC2, y=PC3, color=sex)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by sex") + theme(panel.background = element_blank())

pca.plot.pos.fc1 = ggplot(data=pca.pos.mat.fc,aes(x=PC1, y=PC2, color=sex)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by sex") + theme(panel.background = element_blank())
pca.plot.pos.fc2 = ggplot(data=pca.pos.mat.fc,aes(x=PC1, y=PC3, color=sex)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by sex") + theme(panel.background = element_blank())
pca.plot.pos.fc3 = ggplot(data=pca.pos.mat.fc,aes(x=PC2, y=PC3, color=sex)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by sex") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg.fc1, pca.plot.neg.fc2, pca.plot.neg.fc3, pca.plot.pos.fc1, pca.plot.pos.fc2, pca.plot.pos.fc3, ncol=3)

#by order
table(info.all$order)

pca.neg.mat$order = info.all$order
pca.pos.mat$order = info.all$order

pca.plot.neg2 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=order, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by order") + theme(panel.background = element_blank())

pca.plot.pos2 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=order, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by order") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg2, pca.plot.pos2, ncol=2)

#by bmi
info.all$bmi_c = NULL
info.all$bmi_c = ifelse(info.all$bmi>29.9, 1, 0)
info.all$bmi_c = factor(info.all$bmi_c, levels = c(0,1), labels = c("overweight", "obese"))
table(info.all$bmi_c)

pca.neg.mat$bmi_c = info.all$bmi_c
pca.pos.mat$bmi_c = info.all$bmi_c

pca.plot.neg3 = ggplot(data=pca.neg.mat,aes(x=PC1, y=PC2, color=bmi_c, shape=case)) + geom_point(size=3) + ggtitle("PCA for negative metabolites by BMI") + theme(panel.background = element_blank())

pca.plot.pos3 = ggplot(data=pca.pos.mat,aes(x=PC1, y=PC2, color=bmi_c, shape=case)) + geom_point(size=3) + ggtitle("PCA for positive metabolites by BMI") + theme(panel.background = element_blank())

grid.arrange(pca.plot.neg3, pca.plot.pos3, ncol=2)

#save RData
save(demographics, sample.QC, ID.marker, neg.im, pos.im, neg.wilcox, pos.wilcox, neg.result, pos.result, info.all, pca.pos.mat, pca.neg.mat, file="Glance_afterWilcox_PCA.RData")

```